Subset coercions in \Coq \\ \small{Draft - \today}

### 2006-09-01 papers/article/main.tex
Various fixes
Rework questionable para

The reductions rules of this system are the standard $\beta$ and $\pi$
for projections. We denote by $\downarrow{x}$ the complete normal form
of $x$ and by $\hnf{x}$ its head normal form. We use judgemental equality $`G \type T \eqbp U : s$ and the usual untyped
convertibility $\eqbp$ interchangably (we proved $`G \type T \eqbp U : s
"<=>" T \eqbp U `^ `G \type T, U : s$).

@@

In \Russell, the conversion rule is replaced by a
new subsumption rule \irule{Subsum} (figure
\vref{fig:typing-decl-rules}) which will implement the subset
equivalence. The judgement $`G \typed T \subd U : s$ means $T$ is
equivalent to $U$ in environment $`G$, both sorted with $s$.
Subset equivalence will automatically insert first projections
(\irule{Subset-$\eltpit$}, \irule{SubSub}) and
element introductions (\irule{Element}, \irule{SubProof}), so we do not need
nor want these typing rules in \Russell. We do not want the second
projection to appear in \Russell terms either, because it would lead to
inconsistency. Indeed, in \PVS{} you can't manipulate proof terms, so you have no way to create a
term of type $0 \neq 0$ from the judgement $0 : \mysubset{x}{`N}{x \neq
  0}$, nor can you actually derive the latter, as it won't type-check.
If we allowed the second projection in \Russell, that would be doable.
Hence we leave only the subset type forming rule (figure
\vref{fig:typing-decl-rules}) in \Russell. With this restrictions, we
have the following property:
\begin{proposition}[Subset erasure and consistency]
  If we erase subset types from \Russell terms, leaving only the supports, we get a valid
  \CIC term, hence $`_$ is not provable in \Russell.
\end{proposition}

@@

\end{paragraph}
\begin{paragraph}{}
  The next thing to do is to construct a decidable judgement for
  equivalence given two types. We denote it by $`G \typea T
  \suba U : s$ (figure \ref{fig:subtyping-algo-rules}).
  In our case, we have to check that conversion is decidable, which is the
  case. We also need to eliminate the transitivity and symmetry rules which are not
  syntax-directed. That is the purpose of the following theorem:
\begin{theorem}[Admissibility of transitivity and symmetry]
  \begin{enumerate}
  \item If $`G \typea T \suba U : s$ and $`G \typea U \suba V : s$, then $`G \typea T
    \suba V : s$.
  \item If $`G \typea T \suba U : s$ then $`G \typea U \suba T : s$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  \begin{enumerate}
  \item By induction on the sum of the depths of the two derivations.
  \item By induction, using a thinning lemma.
  \end{enumerate}
\end{proof}
\end{paragraph}

@@

Practically, this means we can use almost all existing definitions in the
\Coq environment. This map is involutive and
the identity on \Russell terms as they do not contain the constructor
$\id{elt}$ or subset projections. Moreover, this erasure function will be
defined on all \Coq terms elaborated from \Russell.
Indeed, the second projection of a
subset element will only appear in the second component of a subset
element. Otherwise, we would have used the projection directly in the
\Russell{} term, where it is forbidden.

>>>>

The reductions rules of this system are the standard $\beta$ and $\pi$
for projections. We denote by $\downarrow{x}$ the complete normal form
of $x$ and by $\hnf{x}$ its head normal form. We use a standard
judgemental equality $`G \type T \eqbp U : s$.

@@

In \Russell, the conversion rule \irule{Conv} is replaced by a
new subsumption rule \irule{Subsum} (figure
\vref{fig:typing-decl-rules}) which will implement the subset
equivalence. The judgement $`G \typed T \subd U : s$ means $T$ is
equivalent to $U$ in environment $`G$, both sorted with $s$.
The essence of our equivalence is to identify subset types if they have
equivalent supports, hence we have the following property:
\begin{proposition}[Subset erasure and consistency]
  If we erase subset types from \Russell terms and rules, leaving only the supports, we get a valid
  \CIC term and derivation, hence $`_$ is not provable in \Russell.
\end{proposition}
\begin{proof}
  By erasing subsets in \Russell rules, the \irule{Subset} rule becomes
  admissible and the \irule{Subsum} rule becomes \irule{Conv}. By
  eliminating the subset type in a term we get a valid \CIC term. Hence
  to each \Russell derivation corresponds a \CIC derivation.
\end{proof}

@@

\end{paragraph}
\begin{paragraph}{}
  The next thing to do is to construct a decidable judgement for
  equivalence given two types. We denote it by $`G \typea T
  \suba U : s$ (figure \ref{fig:subtyping-algo-rules}).
  In our case, we have to check that conversion is decidable, which is the
  case as we can replace judgemental equality with the usual untyped
  convertibility $\eqbp$ (we proved $`G \type T \eqbp U : s
  "<=>" T \eqbp U `^ `G \type T, U : s$).
  We also need to eliminate the transitivity and symmetry rules which are not
  syntax-directed. That is the purpose of the following theorem:
\begin{theorem}[Admissibility of transitivity and symmetry]
  \begin{enumerate}
  \item If $`G \typea T \suba U : s$ and $`G \typea U \suba V : s$, then $`G \typea T
    \suba V : s$.
  \item If $`G \typea T \suba U : s$ then $`G \typea U \suba T : s$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  \begin{enumerate}
  \item By induction on the sum of the depths of the two derivations.
  \item By induction, using a thinning lemma.
  \end{enumerate}
\end{proof}
\end{paragraph}

@@

Practically, this means we can use almost all existing definitions in the
\Coq environment. This map is involutive and
the identity on \Russell terms as they do not contain the constructor
$\id{elt}$ or subset projections. Moreover, this erasure function will be
defined on all \Coq terms elaborated from \Russell.
Indeed, the second projection of a
subset element will only appear in the second component of a subset
element. Otherwise, we would have used the projection directly in the
\Russell{} term, where it is forbidden.
Indeed, in \PVS{} you can't manipulate proof terms, so you have no way to create a
term of type $0 \neq 0$ from the judgement $0 : \mysubset{x}{`N}{x \neq
  0}$, nor can you actually derive the latter, as it won't type-check.
If we allowed the second projection in \Russell, that would be doable
and we would have an inconsistency.
Hence we left only the subset type forming rule (figure
\vref{fig:typing-decl-rules}) in \Russell, while the introduction and
first projection of subset elements are internalized by the coercion judgement.
